
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
#library(fipriskanalysismexico)

#system("sudo apt-get install -y fftw3", wait = TRUE)
#system("sudo apt-get remove libfftw3-dev", wait = TRUE)
#BiocManager::install("EBImage")
devtools::load_all()
```


Estimate statistics from changes in distribution by species and future scenario
```{r sdm statistics, include=FALSE}

sheetlink = "https://docs.google.com/spreadsheets/d/1HoWV3HponuSLKvF1UyBOFXD-gbdwtiuhdui7OUnybZ4"

gs4_deauth()
smartfish.species <- googlesheets4::read_sheet(ss=sheetlink, sheet = "especies_smartfish", col_types = c("cciiiiic"))

#estimate changes in distribution

sdm.scenarios <- c("ssp126", "ssp245", "ssp585")

sdm.species <- tolower(smartfish.species$nombre_cientifico)

sdm.files <- list.files(here::here("data-raw","sdm"), pattern="*.asc", recursive = TRUE)

sdm.names <- tolower(sdm.files)

species.no <- 1:length(sdm.species)
#species.no <- 44:length(species)

crs.eqap <- enmSdmX::getCRS('EE Americas') # Equal area projection Americas, use getCRS() to see all projections available

dist.stats <- lapply(species.no, calc_distchange, sdm.species, crs.eqap, sdm.names, sdm.files, present.name = c("historico","current"), sdm.scenarios) 
  
 
dist.table <- dist.stats %>% bind_rows()
  
list_res <- list()
  
for(eachscenario in 1:length(unique(dist.table$scenario))){

  print(eachscenario)
  this.scenario <- unique(dist.table$scenario)[eachscenario]
  print(this.scenario)
  
  numeric.table <- dist.table %>%
    dplyr::filter(scenario==this.scenario)
  
  process.num <- numeric.table %>% 
    dplyr::select(dplyr::where(is.numeric)) %>%
    caret::preProcess(., method=c("range"))
  
  num.scale <- numeric.table %>%
    dplyr::select(dplyr::where(is.numeric)) %>%
    stats::predict(process.num, .)
  
  dist.table.scaled <- dist.table %>%
    dplyr::filter(scenario==this.scenario) %>%
    dplyr::select(-dplyr::where(is.numeric)) %>%
    dplyr::bind_cols(., num.scale)

  list_res[[eachscenario]] <- dist.table.scaled  
  
}

list_res.table <- do.call(rbind, list_res)

readr::write_csv(list_res.table, here::here("outputs","distribution_stats_table.csv"))

```



=

Get data for components 
```{r get fip data, message=FALSE, warning=FALSE, include=FALSE}

gs4_deauth()

#get species by coop
coop.species <- googlesheets4::read_sheet(ss="https://docs.google.com/spreadsheets/d/1tgKIyHXBG5IQXX2SvN8uTbxvl3nefBqDuQDAw0emFP8/edit?gid=784454269#gid=784454269", sheet="especies_cooperativas") %>%  dplyr::mutate(nombre_comun = tolower(nombre_comun), nombre_cientifico = tolower(nombre_cientifico), cooperativa=tolower(cooperativa))

sheetlink = "https://docs.google.com/spreadsheets/d/1bBn7nLtMHPUSuSWU1XI85AR53cUfr9rSjv_6Dp4OywI"

#get species list
smartfish.species <- googlesheets4::read_sheet(ss=sheetlink, sheet = "especies_smartfish", col_types = c("ccccc")) %>% 
  dplyr::mutate(nombre_comun = tolower(nombre_comun), nombre_cientifico = tolower(nombre_cientifico))

googlesheets4::gs4_deauth()
locs.pesca <- googlesheets4::read_sheet("1zPWnQLfMismgHFkQqmNE8yoLTg3L_3nC24uCTj780nQ", sheet = "localidades", col_types = c("cccccccc")) %>% 
  dplyr::rename(coop_name = cooperativa, cooperativa=nombre_cooperativa) %>% 
  mutate(cooperativa=tolower(cooperativa))
  
speciescoop <- coop.species %>% 
  dplyr::mutate(cooperativa=tolower(cooperativa)) %>%
  dplyr::left_join(smartfish.species, by=c("nombre_cientifico","nombre_comun")) %>% 
  dplyr::filter(!is.na(region)) %>% 
  dplyr::select(-entidad, -localidad) %>% 
  dplyr::left_join(locs.pesca, by="cooperativa")
  

usethis::use_data(speciescoop, overwrite = TRUE)
#creates template describing the data
data.description <-sinew::makeOxygen(speciescoop)
cat(data.description, file=here::here("R",paste0("data-","speciescoop.R")))

#used to check for missing species
#coop.loc.sp %>% dplyr::filter(is.na(region)) %>% View()

#get life history data for species
lifehistorydata <- read_googlesheet(sheetlink, "datos_caracteristicas_smartfish", "ccccccccccccccccccccccccciiiccccccccccccc")
usethis::use_data(lifehistorydata, overwrite = TRUE)
#creates template describing the data
data.description <-sinew::makeOxygen(lifehistorydata)
cat(data.description, file=here::here("R",paste0("data-","lifehistorydata.R")))

lifehistorytraits <- read_googlesheet(sheetlink, "carac_criterios", NULL)
usethis::use_data(lifehistory.traits, overwrite = TRUE)
#creates template describing the data
data.description <-sinew::makeOxygen(lifehistorytraits)
cat(data.description, file=here::here("R",paste0("data-","lifehistorytraits.R")))

#indicators were produced in Packard coastalvulnerabilityproject
adaptivecapacitydata <- readr::read_csv(here::here("data-raw","tot_adaptive capacity.csv"))
usethis::use_data(adaptivecapacitydata, overwrite = TRUE)
#creates template describing the data
data.description <-sinew::makeOxygen(adaptivecapacitydata)
cat(data.description, file=here::here("R",paste0("data-","adaptivecapacitydata.R")))

sensitivitydata <- readr::read_csv(here::here("data-raw","tot_sensitivity.csv"))
usethis::use_data(sensitivitydata, overwrite = TRUE)
#creates template describing the data
data.description <-sinew::makeOxygen(sensitivitydata)
cat(data.description, file=here::here("R",paste0("data-","sensitivitydata.R")))

exposurehistorical <- readr::read_csv(here::here("data-raw","tot_exposure_historical.csv")) 
usethis::use_data(exposurehistorical, overwrite = TRUE)
#creates template describing the data
data.description <-sinew::makeOxygen(exposurehistorical)
cat(data.description, file=here::here("R",paste0("data-","exposurehistorical.R")))

exposuressp126 <- readr::read_csv(here::here("data-raw","tot_exposure_ssp126.csv")) 
usethis::use_data(exposuressp126, overwrite = TRUE)
#creates template describing the data
data.description <-sinew::makeOxygen(exposuressp126)
cat(data.description, file=here::here("R",paste0("data-","exposuressp126.R")))

exposuressp245 <- readr::read_csv(here::here("data-raw","tot_exposure_ssp245.csv")) 
usethis::use_data(exposuressp245, overwrite = TRUE)
#creates template describing the data
data.description <-sinew::makeOxygen(exposuressp245)
cat(data.description, file=here::here("R",paste0("data-","exposuressp245.R")))

exposuressp585 <- readr::read_csv(here::here("data-raw","tot_exposure_ssp585.csv")) 
usethis::use_data(exposuressp585, overwrite = TRUE)
#creates template describing the data
data.description <-sinew::makeOxygen(exposuressp585)
cat(data.description, file=here::here("R",paste0("data-","exposuressp585.R")))

#updates package documentation
devtools::document()

#rebuild and load the package
devtools::load_all() # restarts and loads

 
```


Extract and map locations

```{r get locations}

googlesheets4::gs4_deauth()
locs.pesca <- googlesheets4::read_sheet("1zPWnQLfMismgHFkQqmNE8yoLTg3L_3nC24uCTj780nQ", sheet = "localidades", col_types = c("cccccccc"))

locs.selecc <- unique(locs.pesca$CVE_LOC)

#only communities considered in the analysis
localities  <- readr::read_csv(here::here("data-raw","inegi_coast_comm.csv")) %>% 
  dplyr::filter(CVE_LOC %in% locs.selecc)

usethis::use_data(localities, overwrite = TRUE)
#creates template describing the data
data.description <-sinew::makeOxygen(localities)
cat(data.description, file=here::here("R",paste0("data-","localities.R")))

#https://epsg.io/4485
crs.proj.utm <- ("+proj=utm +zone=12 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs +datum=NAD83") 
crs.proj.wgs <- "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"

#convertir en capa geografica
inegi.2020.coords <- localities %>% 
  dplyr::select("CVE_ENT","CVE_LOC","CVE_MUN","NOM_ENT","NOM_MUN","NOM_LOC","deci_lon","deci_lat") %>% 
  dplyr::left_join(locs.pesca, by=c("CVE_LOC", "CVE_ENT", "CVE_MUN"))

readr::write_csv(inegi.2020.coords, here::here("outputs","smartfish_locations.csv"))

coordinates(inegi.2020.coords) <- c("deci_lon","deci_lat")

#definir proyeccion geografica y proyectar a utm
proj4string(inegi.2020.coords) <- crs.proj.wgs
inegi.2020.coords.sf <- sf::st_as_sf(inegi.2020.coords)
comunidades.coords.proj <- sf::st_transform(inegi.2020.coords.sf, crs.proj.utm)

comunidades.coords.proj %>% 
  sf::st_write(here::here("outputs","shp_files","comunidades_costeras_SmartFish.shp"), append = FALSE)

cost.utm.buff <- sf::st_buffer(comunidades.coords.proj, dist=10000)

cost.utm.buff %>% 
  sf::st_write(here("outputs","shp_files","comunidades_costeras_SmartFish_buffer.shp"), append = FALSE)

#comunidades.coords.proj  <-  
#  sf::st_read(here::here("outputs","shp_files","comunidades_costeras_SmartFish.shp"), append = FALSE)

coastal.states <- c("Baja California","Baja California Sur","Nayarit")

coast.state.mx <- sf::st_read(here::here("data-raw","shapefiles","ESTADOS.shp")) %>% 
                       sf::st_transform(6362) %>% 
  dplyr::filter(NOM_ENT %in% coastal.states)

map.mx <- sf::st_read(here::here("data-raw","shapefiles","Mexico_lm.shp"))%>% 
                       sf::st_transform(6362) 

coast_points <- sf::st_centroid(coast.state.mx)
coast_points.cor <- cbind(coast.state.mx, st_coordinates(sf::st_centroid(coast.state.mx$geometry)))

#plot

sf::st_bbox(map.mx)

subs.map <- ggplot2::ggplot(data = coast.state.mx) +
  ggplot2::geom_sf() +
  ggplot2::geom_text(data= coast_points.cor,ggplot2::aes(x=X, y=Y, label=NOM_ENT), color = "black", check_overlap = TRUE, size = 3) +
  ggplot2::xlab("Longitude") + ggplot2::ylab("Latitude") +
  ggspatial::annotation_scale(location = "bl", width_hint = 0.5) +
  ggspatial::annotation_north_arrow(location = "tr", which_north = "true", 
                                    pad_x = unit(0.2, "in"), 
                                    pad_y = unit(0.2, "in")) +
  ggplot2::theme_light() + 
  ggplot2::geom_sf(data = map.mx, fill = "transparent", color = "black") +
  ggplot2::geom_sf(data = comunidades.coords.proj, mapping = ggplot2::aes(color = cooperativa), size=2) +
  ggplot2::scale_color_manual(values = RColorBrewer::brewer.pal(8, "Set1")) +
  ggplot2::labs(x= "Lon", y= "Lat", color = "Cooperativa") +
  coord_sf(xlim = c(2581774.6, 1070890.3), ylim = c(2349711.9, 821154.9)) +
  theme(legend.position="bottom") +
  ggplot2::guides(colour=guide_legend(nrow=2))

ggplot2::ggsave(here::here("outputs","figures","mapa_localidades_smartfish.png"), plot = subs.map, width = 19, height = 14, units = "cm", dpi = 600)


```


Estimate vulnerability, for its different components
```{r get vulnerability}

# coms.cost <- readr::read_csv(here::here("data-raw","comunidades_costeras_INEGI.csv"))
#get data community vulnerability
data("adaptivecapacitydata")

dir.create(here::here("outputs"))
dir.create(here::here("outputs","analysis"))
dir.create(here::here("outputs","figures"))

sm.locations <- readr::read_csv(here::here("outputs","smartfish_locations.csv"))

fiplocal.id <- unique(sm.locations$CVE_LOC)

googlesheets4::gs4_deauth()
locs.pesca <- googlesheets4::read_sheet("1zPWnQLfMismgHFkQqmNE8yoLTg3L_3nC24uCTj780nQ", sheet = "localidades", col_types = c("cccccccc"))


#variables utilizadas

ind.vul<- googlesheets4::read_sheet("1zPWnQLfMismgHFkQqmNE8yoLTg3L_3nC24uCTj780nQ", sheet = "indicadores") %>% 
  dplyr::filter(component=="Vulnerabilidad") %>% 
  dplyr::distinct(variable_id, direction)

vul.ind.names <- ind.vul$variable_id

adaptive.bc <- adaptivecapacitydata %>% 
  dplyr::select(CVE_LOC, all_of(vul.ind.names)) %>%
  dplyr::filter(CVE_LOC %in% fiplocal.id) %>% 
  tidyr::pivot_longer(cols=2:(ncol(.)), names_to="variable_id",values_to="value") 

#test for lack of variation between variables

adaptive.bc.var <- adaptive.bc %>% 
  group_by(variable_id) %>%
  summarise(range=max(value)-min(value))

test.var  <-  adaptive.bc.var %>% 
  dplyr::filter(range==0)  

vul.vars <- adaptive.bc %>% 
  dplyr::filter(!variable_id %in% test.var$variable_id) %>%
  dplyr::mutate(value = dplyr::coalesce(value, 0)) %>% 
  pivot_wider(names_from = variable_id, values_from = value) 

process.vul <- vul.vars %>%
  dplyr::select(dplyr::where(is.numeric)) %>%
  caret::preProcess(., method=c("range"))

#correct directionality so that higher is negative

vul.data.set.sc <- vul.vars %>% 
  dplyr::select(dplyr::where(is.numeric)) %>%
  stats::predict(process.vul, .) %>% 
  dplyr::bind_cols(CVE_LOC=vul.vars$CVE_LOC, .) %>% 
  tidyr::pivot_longer(cols=2:ncol(.), names_to="variable_id",values_to="value") %>% 
  dplyr::left_join(ind.vul, by="variable_id") %>% 
  mutate(value = ifelse(direction == "higher_positive", 1 - value, value)) %>% 
  dplyr::select(-direction)


#get data for cooperatives

googlesheets4::gs4_deauth()
ind.coop <- googlesheets4::read_sheet("1zPWnQLfMismgHFkQqmNE8yoLTg3L_3nC24uCTj780nQ", sheet = "indicadores_cooperativa", col_types = c("ccccd")) 

ind.catego <- googlesheets4::read_sheet("1zPWnQLfMismgHFkQqmNE8yoLTg3L_3nC24uCTj780nQ", sheet = "catego_ind_coop", col_types = c("cccc")) 

ind.coop.wide <- ind.coop %>%
  dplyr::select(-indicador_categoria, -indicador) %>%
  tidyr::pivot_wider(names_from = nom_abrev, values_from = valor)  

process.coop <- ind.coop.wide %>% 
  dplyr::select(dplyr::where(is.numeric)) %>%
  caret::preProcess(., method=c("range"))

vul.data.set.coop <- ind.coop.wide %>% 
  dplyr::select(dplyr::where(is.numeric)) %>%
  stats::predict(process.coop, .) %>% 
  dplyr::bind_cols(cooperativa=ind.coop.wide$cooperativa, .) %>% 
  pivot_longer(cols=2:ncol(.), names_to="indicador",values_to="value") %>%
  dplyr::left_join(ind.catego, by="indicador") %>%
  mutate(value = ifelse(direction == "higher_positive", 1 - value, value)) %>% 
  dplyr::select(-indicador_categoria, -componente, -direction) %>% 
  dplyr::rename(variable_id = indicador)


vul.data.set.wide <- vul.data.set.sc %>% 
 tidyr::pivot_wider(names_from = variable_id, values_from = value)


scaled.vul <- vul.data.set.coop %>% 
  dplyr::left_join(locs.pesca, by ="cooperativa") %>%
  tidyr::pivot_wider(names_from = variable_id, values_from = value) %>% 
  dplyr::left_join(vul.data.set.wide, by=c("CVE_LOC"))

#use to find duplicates
#vul.data.set.coop |>
#  dplyr::summarise(n = dplyr::n(), .by = c(CVE_LOC, variable_id)) |>
#  dplyr::filter(n > 1L)

ind.vul <- googlesheets4::read_sheet("1zPWnQLfMismgHFkQqmNE8yoLTg3L_3nC24uCTj780nQ", sheet = "indicadores") %>% 
  dplyr::filter(component=="Vulnerabilidad") %>% 
  dplyr::distinct(variable_id, Indicator_category) %>% 
  dplyr::rename(nom_abrev = variable_id, indicador_categoria = Indicator_category)

cat.vul <- ind.coop %>% 
  dplyr::distinct(indicador_categoria, nom_abrev) %>% 
  dplyr::bind_rows(ind.vul)

scaled.vul.long <- scaled.vul %>% 
  tidyr::pivot_longer(cols=9:ncol(.), names_to="nom_abrev",values_to="value") %>% 
  dplyr::left_join(cat.vul, by="nom_abrev") 

readr::write_csv(scaled.vul.long, here::here("outputs","Tabla_vulnerabilidad.csv"))

readr::write_csv(scaled.vul.long, here::here("outputs","scaled_vulnerability.csv"))

```

Estimate exposure

```{r Exposure}
data("adaptivecapacitydata")
data("sensitivitydata")
data("localities")

indicator.data <- adaptivecapacitydata %>% 
  dplyr::left_join(sensitivitydata, by=c("CVE_LOC","CVE_MUN"))

sm.locations <- readr::read_csv(here::here("outputs","smartfish_locations.csv"))

fiplocal.id <- unique(sm.locations$CVE_LOC)

googlesheets4::gs4_deauth()
locs.pesca <- googlesheets4::read_sheet("1zPWnQLfMismgHFkQqmNE8yoLTg3L_3nC24uCTj780nQ", sheet = "localidades", col_types = c("cccccccc"))

caracs.pesq <- googlesheets4::read_sheet("1zPWnQLfMismgHFkQqmNE8yoLTg3L_3nC24uCTj780nQ", sheet = "indicadores_pesquerias")

ind.expo <- googlesheets4::read_sheet("1zPWnQLfMismgHFkQqmNE8yoLTg3L_3nC24uCTj780nQ", sheet = "indicadores") %>% 
  dplyr::filter(component=="Exposicion") %>% 
  dplyr::distinct(variable_id, direction)

#fisheries structure
expo.ind.names <- ind.expo$variable_id

exposure.bc <- indicator.data %>% 
  dplyr::select(CVE_LOC, all_of(expo.ind.names)) %>%
  dplyr::filter(CVE_LOC %in% fiplocal.id) %>% 
  mutate_if(is.numeric, funs(ifelse(is.na(.), 0, .))) 

process.coop <- exposure.bc %>% 
  dplyr::select(dplyr::where(is.numeric)) %>%
  caret::preProcess(., method=c("range"))

expo.data.set.coop <- exposure.bc %>% 
  dplyr::select(dplyr::where(is.numeric)) %>%
  stats::predict(process.coop, .) %>% 
  dplyr::bind_cols(CVE_LOC=exposure.bc$CVE_LOC, .) 
 

#fisheries characteristics

ind.coop.wide <- caracs.pesq %>%
  dplyr::select(cooperativa, nom_abrev, valor) %>%
  tidyr::pivot_wider(names_from = nom_abrev, values_from = valor)  

process.coop <- ind.coop.wide %>% 
  dplyr::select(dplyr::where(is.numeric)) %>%
  caret::preProcess(., method=c("range"))

expo.data.scaled <- ind.coop.wide %>% 
  dplyr::select(dplyr::where(is.numeric)) %>%
  stats::predict(process.coop, .) %>% 
  dplyr::bind_cols(cooperativa=ind.coop.wide$cooperativa, .) 
 

ind.cat <- googlesheets4::read_sheet("1zPWnQLfMismgHFkQqmNE8yoLTg3L_3nC24uCTj780nQ", sheet = "indicadores") %>% 
  dplyr::filter(component=="Exposicion") %>% 
  dplyr::distinct(variable_id, Indicator_category) %>% 
  dplyr::rename(indicator = variable_id, indicador_categoria = Indicator_category)

carac.cat <- caracs.pesq %>% 
  dplyr::distinct(indicador_categoria, nom_abrev) %>% 
  dplyr::rename(indicator = nom_abrev) %>% 
  dplyr::bind_rows(ind.cat)

ind.dir <- ind.expo %>% 
  dplyr::rename(indicator=variable_id)

# corrected directionality, higher is negative
scaled.expo <- expo.data.scaled %>% 
  dplyr::left_join(locs.pesca, by ="cooperativa") %>%
  dplyr::left_join(expo.data.set.coop, by=c("CVE_LOC")) %>% 
  dplyr::select(cooperativa, nombre_cooperativa, localidad, entidad, mun, CVE_ENT, CVE_MUN, CVE_LOC, everything()) %>% 
  tidyr::pivot_longer(cols=9:ncol(.), names_to="indicator",values_to="value") %>%
  dplyr::left_join(carac.cat, by="indicator") %>% 
  dplyr::left_join(ind.dir, by="indicator") %>%
  dplyr::mutate(direction = ifelse(is.na(direction), "higher_positive", direction)) %>% 
  dplyr::mutate(value = ifelse(direction == "higher_positive", 1 - value, value)) %>% 
  dplyr::select(-direction) %>% 
  dplyr::mutate(value=ifelse(is.nan(value), 0, value))


readr::write_csv(scaled.expo,here::here("outputs","scaled_exposure.csv"))  

```

Obtain life history data by species

```{r load data, message=FALSE, warning=FALSE, include=FALSE}

data("fipinventory")
data("lifehistorydata")
data("localities")
data("lifehistorytraits")

species.sm <- googlesheets4::read_sheet("1bBn7nLtMHPUSuSWU1XI85AR53cUfr9rSjv_6Dp4OywI", sheet = "especies_smartfish")

lifehistory.bc <- get_lifehistory(lifehistorydata, localities, lifehistorytraits, species.sm)

lifehistory.bc %>% View()

readr::write_csv(lifehistory.bc, here::here("outputs","lifehistory_data.csv"))


```



Estimate species sensitivity

```{r species sensitivity}


lifehistory.trait <- readr::read_csv(here::here("outputs","lifehistory_data.csv")) %>% 
  dplyr::select(-value, -Type, -category, -status, -topsis_weight) %>% 
  tidyr::pivot_wider(names_from = traits, values_from = score)

lifehistory <- lifehistory.trait %>% 
  dplyr::select(-species) %>% 
  names()
  
scaled.lifehistory <- scale_var(datatable=lifehistory.trait, sspscenario=NA)

readr::write_csv(scaled.lifehistory, here::here("outputs","sensibilidad_historiavida.csv"))

readr::write_csv(scaled.lifehistory, here::here("outputs","scaled_lifehistory.csv"))

```



Estimate hazard
```{r get hazard}

data("exposurehistorical")
data("exposuressp126")
data("exposuressp245")
data("exposuressp585")
data("localities")

#exposure by communities

googlesheets4::gs4_deauth()
vars.expo <- googlesheets4::read_sheet("1eJXjwlHqEzSAZaTmE_XyN93fu6EI0I8igRH4AdvXOl8", sheet="vars_updated")

fiplocal.id <- localities %>%
    dplyr::distinct(CVE_LOC) %>% 
    dplyr::pull(CVE_LOC)

expohist.bc <- exposurehistorical %>% 
  dplyr::filter(CVE_LOC %in% fiplocal.id) 


scaled.expo.hist <- scale_var_expo(datatable=exposurehistorical, sspscenario="historical", vars.expo, fiplocal.id) %>% 
  tidyr::pivot_longer(cols=3:ncol(.), names_to="indicator",values_to="value")
  
scaled.expo.ssp126 <- scale_var_expo(datatable=exposuressp126, sspscenario="ssp126", vars.expo, fiplocal.id)%>% 
  tidyr::pivot_longer(cols=3:ncol(.), names_to="indicator",values_to="value")

scaled.expo.ssp245 <- scale_var_expo(datatable=exposuressp126, sspscenario="ssp245", vars.expo, fiplocal.id)%>% 
  tidyr::pivot_longer(cols=3:ncol(.), names_to="indicator",values_to="value")

scaled.expo.ssp585 <- scale_var_expo(datatable=exposuressp126, sspscenario="ssp585", vars.expo, fiplocal.id)%>% 
  tidyr::pivot_longer(cols=3:ncol(.), names_to="indicator",values_to="value")


googlesheets4::gs4_deauth()
locs.pesca <- googlesheets4::read_sheet("1zPWnQLfMismgHFkQqmNE8yoLTg3L_3nC24uCTj780nQ", sheet = "localidades", col_types = c("cccccccc"))

scaled.hazard <- dplyr::bind_rows(scaled.expo.hist, scaled.expo.ssp126, scaled.expo.ssp245, scaled.expo.ssp585) %>% 
  dplyr::left_join(locs.pesca, by="CVE_LOC") %>% 
  dplyr::mutate(component="hazard")

readr::write_csv(scaled.hazard, here::here("outputs","scaled_hazard.csv"))

```

plot sensitivity 
```{r}
data("speciescoop")

#lifehistory

sens.lifehistory <- readr::read_csv(here::here("outputs","scaled_lifehistory.csv")) %>% 
  dplyr::mutate(species = tolower(species))
print(sens.lifehistory)

sp.coop.nom <- speciescoop %>% 
  dplyr::rename(species = nombre_cientifico)

sens.coop.lf <- sp.coop.nom %>%
  dplyr::left_join(sens.lifehistory, by="species") %>% 
  dplyr::filter(!is.na(tot_score)) %>% 
  dplyr::select(-nom_coop, -arte_pesca, -fecha_fip, -lider_fip, -nombre_fip, -region)

gs4_deauth()
#fix directionality of life history attributes, some were eliminated because it was unclear they could affect response to climate change
vars.lf <- googlesheets4::read_sheet("1bBn7nLtMHPUSuSWU1XI85AR53cUfr9rSjv_6Dp4OywI", sheet="life_history_direction")

names(sens.lifehistory)

#check for missing species
#sens.coop.lf %>% dplyr::filter(is.na(tot_score)) %>% View()
  
sens.lifehistory.long <- sens.coop.lf %>% 
  tidyr::pivot_longer(cols = 14:ncol(.), names_to = "var_name", values_to = "value") %>% 
  dplyr::left_join(vars.lf, by="var_name") %>% 
  dplyr::filter(!is.na(direction)) %>% 
  dplyr::mutate(value = dplyr::if_else(direction == "higher_positive", 1 - value, value)) %>% 
  dplyr::select(-direction, -var_name) %>% 
  dplyr::rename(var_name = var_name_sp)

print(sens.lifehistory.long)

sens.lifehistory.natest <- sens.lifehistory.long %>% 
  dplyr::filter(is.na(value))

sens.lf.long.mean <- sens.lifehistory.long %>% 
  dplyr::group_by(cooperativa, coop_name, localidad, habitat_group, status, entidad, species) %>% 
  dplyr::summarise(mean_value = mean(value))

#sp.cols <- c("#E69F00", "#56B4E9", "#873e23", "#F0E442", "#0072B2", "#999999")


sens.lf.plot <- sens.lf.long.mean %>% 
  dplyr::mutate(Estatus = dplyr::if_else(status=="fip","FIP", "Otros")) %>% 
#  dplyr::mutate(species=str_to_sentence(species)) %>% 
  ggplot2::ggplot(ggplot2::aes(x=habitat_group, y=mean_value)) + 
  ggplot2::geom_point(ggplot2::aes(color = Estatus),
    size = 3, 
    alpha = 0.7 # It's nice to add some transparency because there may be overlap.
  ) +
 # ggplot2::scale_color_manual(values=sp.cols, name="Especie") +
  ggplot2::facet_grid(coop_name ~ .) +
  ggplot2::scale_color_brewer(palette = "Dark2") %>% 
  ggplot2::labs(
    title = "Sensibilidad relativa de grupos en base a la historia de vida",
    subtitle = "Un valor alto es mayor sensibilidad al cambio climático",
    x = "Grupo",
    y = "Valor del indicador") +
#  ggplot2::theme(legend.position="none")+
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90, vjust = 1, hjust=1, size = 11))+
  ggplot2::theme(strip.text = ggplot2::element_text(
    size = 11))+
  ggplot2::ylim(0,1) +
  ggplot2::theme(legend.position="bottom")+
  ggplot2::theme_linedraw() 

sens.lf.plot



ggplot2::ggsave("sensibilidad_historia_vida.png", sens.lf.plot, path = here::here("outputs","figures"), dpi=400, width= 14, height=14,scale =1)



#measures of biotic velocity
#modified so that high value of sensitivity is negative

sens.dist <- readr::read_csv(here::here("outputs","distribution_stats_table.csv"))  %>% 
  dplyr::mutate_at(c("suitability", "rmsq", "godsoeESp", "schoenerD", "warrenI","cor","rankCor"),funs(1-.))

print(sens.dist)
#centroid velocity is how quickly the species shifts distribution (higher value negative)
#suitability (higher value positive)
#correlation values are correlations with original distribution (higher value positive)
#indices of niche overlap rmsd, godsoeESp, schoenerD, warren I, cor Rank cor measure higher values are positive, less change

sens.data.long <- speciescoop %>% 
  dplyr::rename(species = nombre_cientifico) %>% 
  dplyr::select(-nom_coop, -arte_pesca, -fecha_fip, -lider_fip, -nombre_fip, -region) %>%
  dplyr::left_join(sens.dist, by="species") %>%
  tidyr::pivot_longer(cols = 15:ncol(.), names_to = "var_name", values_to = "value")

sens.data.long.mean <- sens.data.long %>% 
  dplyr::group_by(cooperativa, coop_name, localidad, entidad, species, scenario) %>% 
  dplyr::summarise(mean_value = mean(value)) %>% 
  dplyr::left_join(sp.coop.nom, by=c("coop_name","cooperativa","species","localidad","entidad"))


sens.dist.plot <- sens.data.long.mean %>% 
  dplyr::mutate(Estatus = dplyr::if_else(status=="fip","FIP", "Otros")) %>% 
#  dplyr::mutate(species=str_to_sentence(species)) %>% 
  ggplot2::ggplot(ggplot2::aes(x=habitat_group, y=mean_value)) + 
  ggplot2::geom_point(ggplot2::aes(color = Estatus),
    size = 3, 
    alpha = 0.7 # It's nice to add some transparency because there may be overlap.
  ) +
 # ggplot2::scale_color_manual(values=sp.cols, name="Especie") +
  ggplot2::facet_grid(coop_name ~ scenario) +
  ggplot2::scale_color_brewer(palette = "Dark2") %>% 
  ggplot2::labs(
    title = "Sensibilidad relativa de grupos en base a la distribución",
    subtitle = "Un valor alto es mayor sensibilidad al cambio climático",
    x = "Grupo",
    y = "Valor del indicador") +
#  ggplot2::theme(legend.position="none")+
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90, vjust = 1, hjust=1, size = 11))+
  ggplot2::theme(strip.text = ggplot2::element_text(
    size = 11))+
  ggplot2::ylim(0,1) +
  ggplot2::theme(legend.position="bottom")+
  ggplot2::theme_linedraw() 

sens.dist.plot

ggplot2::ggsave("sensibilidad_vel_biotica.png", sens.dist.plot, path = here::here("outputs","figures"), dpi=400, width= 12, height=14,scale =1)


#joint two measures of sensitivity

sens.all.mean <- sens.lf.long.mean %>% 
  dplyr::rename(mean_value_lf = mean_value) %>%
  dplyr::left_join(sens.data.long.mean, by=c("cooperativa","coop_name","species","localidad","entidad","habitat_group","status")) %>%
  dplyr::rename(mean_value_dist= mean_value)

sens.all.mean %>% readr::write_csv(here::here("outputs","scaled_sensitivity.csv"))  


sens.all.plot <- sens.all.mean %>% 
  dplyr::mutate(Estatus = dplyr::if_else(status=="fip","FIP", "Otros")) %>% 
#  dplyr::mutate(species=str_to_sentence(species)) %>% 
  ggplot2::ggplot(ggplot2::aes(x=mean_value_lf, y=mean_value_dist)) + 
  ggplot2::geom_point(ggplot2::aes(color = habitat_group, shape = Estatus),
    size = 3, 
    alpha = 0.7 # It's nice to add some transparency because there may be overlap.
  ) +
 # ggplot2::scale_color_manual(values=sp.cols, name="Especie") +
  ggplot2::facet_grid(coop_name ~ scenario) +
  ggplot2::scale_color_brewer(palette = "Dark2") %>% 
  ggplot2::labs(
    title = "Sensibilidad de grupos de especies para cada cooperativa",
    subtitle = "Un valor alto es mayor sensibilidad al cambio climático",
    x = "Sensibilidad media en base a la historia de vida",
    y = "Sensibilidad media en base a cambios en la distribución", 
    color = "Grupo") +
#  ggplot2::theme(legend.position="none")+
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90, vjust = 1, hjust=1, size = 11))+
  ggplot2::theme(strip.text = ggplot2::element_text(
    size = 11))+
  ggplot2::ylim(0,1) +
  ggplot2::theme(legend.position="bottom") +
  ggplot2::theme_linedraw() 
                

sens.all.plot

ggplot2::ggsave("sensibilidad_all.png", sens.all.plot, path = here::here("outputs","figures"), dpi=400, width= 12, height=14,scale =1)


```

Plot vulnerability, hazard and exposure

```{r}

sens.data <- readr::read_csv(here::here("outputs","scaled_sensitivity.csv"))  

sens.sc.data.dist <- sens.data %>%
  dplyr::select(-mean_value_lf) %>%
  dplyr::group_by(coop_name, esp_categoria, habitat_group) %>%
  dplyr::summarise(mean_value=mean(mean_value_dist)) %>% 
  dplyr::mutate(componente="sensibilidad hv")
 
sens.sc.data <- sens.data %>%
  dplyr::select(-mean_value_dist) %>%
  dplyr::group_by(coop_name, esp_categoria, habitat_group) %>%
  dplyr::summarise(mean_value=mean(mean_value_lf)) %>% 
  dplyr::mutate(componente="sensibilidad dist")  %>% 
  dplyr::bind_rows(sens.sc.data.dist) %>% 
  dplyr::rename(cooperativa=coop_name, indicador_categoria = esp_categoria, indicator=habitat_group) %>% 
  dplyr::mutate(scenario="historical")

vul.sc.data <-
  readr::read_csv(here::here("outputs", "scaled_vulnerability.csv"))  %>%
  dplyr::mutate(componente = "vulnerabilidad") %>% 
  dplyr::group_by(componente, cooperativa, indicador_categoria, nom_abrev) %>%
  dplyr::summarise(mean_value=mean(value)) %>% 
  dplyr::mutate(scenario="historical") %>% 
  dplyr::rename(indicator = nom_abrev)

expo.sc.data <-
  readr::read_csv(here::here("outputs", "scaled_exposure.csv"))  %>%
  dplyr::mutate(componente = "exposure") %>% 
  dplyr::group_by(componente, cooperativa, indicador_categoria, indicator) %>%
  dplyr::summarise(mean_value=mean(value)) %>% 
  dplyr::mutate(scenario="historical")
  
hazard.sc.data <-
  readr::read_csv(here::here("outputs", "scaled_hazard.csv")) %>% 
  dplyr::rename(componente = component) %>%
  dplyr::filter(indicator!="tot_score") %>% 
  dplyr::group_by(componente, cooperativa, scenario, indicator) %>%
  dplyr::summarise(mean_value=mean(value, na.rm = TRUE)) %>% 
  dplyr::mutate(indicador_categoria="peligro") 
  

risk.data <- dplyr::bind_rows(sens.sc.data, vul.sc.data, expo.sc.data, hazard.sc.data) %>% 
  dplyr::ungroup() %>% 
  dplyr::mutate(indicador_categoria = dplyr::if_else(indicador_categoria=="hazard", "ambientales", indicador_categoria)) %>%
  dplyr::mutate(eje = dplyr::if_else(componente=="sensibilidad dist", "peligro",
                                     dplyr::if_else(componente=="sensibilidad hv", "peligro",
                                            dplyr::if_else(componente=="hazard", "peligro", 
                                                           dplyr::if_else(componente=="exposure", "exposición", componente))))) 

risk.data %>% dplyr::distinct(eje)

  sp.cols <-
  c("#154c79",
             "#76b5c5",
             "#873e23",
             "#0072B2",
             "#56B4E9",
             "#999999")
  
  comp.plot <- risk.data %>%
    dplyr::filter(scenario=="historical") %>% 
    ggplot2::ggplot(ggplot2::aes(x = eje, y = mean_value)) +
    ggplot2::geom_boxplot(ggplot2::aes(color = eje)) +
    ggplot2::geom_jitter(ggplot2::aes(color = eje), alpha = 0.8) +
    ggplot2::facet_wrap(. ~ cooperativa) + # It's nice to add some transparency because there may be overlap.)
    ggplot2::scale_color_manual(values = sp.cols, name = "Componente") +
    ggplot2::labs(
      title = "Componentes del riesgo ante el cambio climático bajo condiciones actuales",
      subtitle = "Los puntos son diferentes indicadores",
      x = "Eje",
      y = "Valor promedio"
    ) +
    ggplot2::theme_linedraw() +
    ggplot2::theme(legend.position="none") +
    ggplot2::theme(strip.text = ggplot2::element_text(
      size = 11))
                     #   ggplot2::facet_wrap(.~FIP, ncol=2) 
                      
ggplot2::ggsave("componentes_riesgo.png", comp.plot, path = here::here("outputs","figures"), dpi=400, width= 16, height=11,scale =1)


```



Plot risk
```{r plot risk}


topsis.res <-
  lapply(hazard.sc, calc_topsis, hazard.sc.data, vul.expo.dat, sp.cols) %>%
  dplyr::bind_rows() %>%
  dplyr::mutate(escenario = dplyr::if_else(escenario == "historico", "histórico", escenario)) %>% left_join(fipinventory, by = c("fip_name")) %>% 
  distinct(escenario, fip_name, fip_id_number, species, score) %>% 
  left_join(specieslist, by="species") 

topsis.res$taxa <- factor(topsis.res$taxa, levels = c("Bivalves","Crabs","Lobster","Shrimp","Sea urchin","Octopus","Reef fish","Demersal fish","Pelagic fish","Pelagic-oceanic fish"))

sp.cols <-
  c("#ffd480","#ffd11a" ,"#E69F00","#ff0000","#ff751a","#800000","#56B4E9","#0072B2","#751aff","#003399")
             
topsis.plot <- topsis.res %>%
  dplyr::mutate(scenario=if_else(escenario=="histórico", "Historical", escenario)) %>% 
  dplyr::filter(!is.na(taxa)) %>% 
  dplyr::filter(scenario=="Historical") %>% 
  dplyr::mutate(fip_id_number = as.factor(fip_id_number)) %>%
  dplyr::arrange(taxa) %>% 
  ggplot2::ggplot(ggplot2::aes(x = taxa, y = score, color = taxa)) +
  ggplot2::geom_point(size = 2,
                      alpha = 0.8) + # It's nice to add some transparency because there may be overlap.)
             #         ggplot2::scale_color_manual(values = sp.cols, name = "fip_name") +
                        ggplot2::labs(
                          title = "Risk to fisheries due to climate change",
                          subtitle = "Higher values indicate more risk",
                          x = "FIP ID number",
                          y = "Taxa"
                        ) +
                        ggplot2::theme_linedraw() +
                        ggplot2::scale_shape(guide = "none") +
                        ggplot2::scale_color_manual(values = sp.cols) +
  theme(axis.text.x = element_text(angle = 45))
                      #  ggplot2::theme(legend.position="none")
                      
                      # # https://stackoverflow.com/questions/60917778/how-do-i-plot-an-image-from-phylopic-in-top-right-corner-of-my-ggplot-graph-in-r
                      # verdillo <-
                      #   png::readPNG(
                      #     RCurl::getURLContent(
                      #       "https://images.phylopic.org/images/eb6f54bf-3182-4d3a-9f16-602759c76db0/raster/512x253.png"
                      #     ),
                      #     native = T
                      #   )
                      # verdillo.grob <- grid::rasterGrob(verdillo, interpolate = TRUE)
                      # 
                      # erizo <-
                      #   png::readPNG(
                      #     RCurl::getURLContent(
                      #       "https://images.phylopic.org/images/0552ec39-a86b-4e23-b92a-7c1c38386e29/raster/1024x715.png"
                      #     ),
                      #     native = T
                      #   )
                      # erizo.grob <- grid::rasterGrob(erizo, interpolate = TRUE)
                      # 
                      # pulpo <-
                      #   png::readPNG(
                      #     RCurl::getURLContent(
                      #       "https://images.phylopic.org/images/f060bcc7-5725-46f7-8276-14553af3707f/raster/1452x1536.png"
                      #     ),
                      #     native = T
                      #   )
                      # pulpo.grob <- grid::rasterGrob(pulpo, interpolate = TRUE)
                      # 
                      # langosta <-
                      #   png::readPNG(
                      #     RCurl::getURLContent(
                      #       "https://images.phylopic.org/images/d96e18f2-c9d6-4b2e-914c-678ed3c72a28/raster/759x1024.png"
                      #     ),
                      #     native = T
                      #   )
                      # langosta.grob <- grid::rasterGrob(langosta, interpolate = TRUE)
ggplot2::ggsave(
  "riesgo_pesqueria.png",
  topsis.plot,
  path = here::here("outputs", "figures"),
  dpi = 400,
  width = 8,
  height = 8,
  scale = 1
)                      # 
                      
topsis.plot.icons <- topsis.plot +
                        ggplot2::annotation_custom(
                          erizo.grob,
                          xmin = 3.5,
                          xmax = 4,
                          ymin = 0.6,
                          ymax = Inf
                        ) +
                        ggplot2::annotation_custom(
                          langosta.grob,
                          xmin = 3.5,
                          xmax = 4,
                          ymin = 0.52,
                          ymax = Inf
                        ) +
                        ggplot2::annotation_custom(
                          pulpo.grob,
                          xmin = 3.5,
                          xmax = 4,
                          ymin = 0.35,
                          ymax = Inf
                        ) +
                        ggplot2::annotation_custom(
                          verdillo.grob,
                          xmin = 3.5,
                          xmax = 4,
                          ymin = 0.05,
                          ymax = Inf
                        )
                      
                      
ggplot2::ggsave(
  "riesgo_pesqueria.png",
  topsis.plot.icons,
  path = here::here("outputs", "figures"),
  dpi = 400,
  width = 8,
  height = 8,
  scale = 1
)
                      

```

